---
title: "Homework 1"
author: "Ani Bilazarian ab4797"
date: "9/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1 

_Creating a Data Frame_
```{r df_create}
library(tidyverse)
set.seed(100)
hw_problem1_df = tibble(
  samp = rnorm(8),
  vec_logical = samp > 0,
  vec_char = c("this", "homework", "is", "kind", "of", "hard", "still", "trying"),
  vec_factor = factor(c("a", "a", "a", "b", "b", "b", "c", "c")))
``` 

_Taking the Means_

* Sample Mean
```{r sample}
mean_samp = mean(pull(hw_problem1_df, samp))
mean_samp
```

* Vec_Logical mean
```{r}
mean_vec_logical = mean(pull(hw_problem1_df, vec_logical))
mean_vec_logical
```

* Vec_Factor & Vec_Character mean

We can only take the mean of a numeric or logical variable. Taking the mean of a factor or character vector results in an error. 

```{r}
mean_vec_char = mean(pull(hw_problem1_df, vec_char))
mean_vec_factor = mean(pull(hw_problem1_df, vec_factor))
```

_Converting Variables: Chunk 1_

```{r results = 'hide', warning = FALSE}
as.numeric(pull(hw_problem1_df,vec_char))
as.numeric(pull(hw_problem1_df,vec_factor))
as.numeric(pull(hw_problem1_df,vec_logical))
```

We can only convert factor and logical vectors to numeric. We cannot convert a character vector to a numeric. 

_Converting Variables: Chunk 2_

* Logic to Numeric and Multiplied by the Result

```{r}
vec_logical_numeric = as.numeric(pull(hw_problem1_df,vec_logical))
(vec_logical_numeric) * (pull(hw_problem1_df,samp))
```

* Logic to Factor and Multiplied by the Result

```{r results = 'hide', warning = FALSE}
vec_logical_factor = as.factor(pull(hw_problem1_df,vec_logical))
(vec_logical_factor) * (pull(hw_problem1_df,samp))
```
Here we can convert the logical vector to a factor but we cannot multiply it by the result as a factor vector cannot be multiplied.  

* Logic to Factor to Numeric and Multiplied by the Result

```{r}
as.numeric(vec_logical_factor) * (pull(hw_problem1_df,samp))
```

# Problem 2 

_Creating a Data Frame_
```{r}
hw_problem2_df = tibble(
  x = rnorm(500, mean = 0, sd = 1 ),
  y = rnorm(500, mean = 0, sd = 1),
  vec_logical = c(x + y > 1 ),
  vec_numeric = as.numeric(vec_logical),
  vec_factor = as.factor(vec_logical)
  )
```

_Data Frame Description_
```{r results = 'hide', warning = FALSE}
nrow(hw_problem2_df)
ncol(hw_problem2_df)
mean(pull(hw_problem2_df, x))
median(pull(hw_problem2_df, x))
sd(pull(hw_problem2_df, x))
sum(pull(hw_problem2_df, vec_logical), na.rm=TRUE)/500
```

In this data frame we can see that the number of rows is `r nrow(hw_problem2_df)`, the number of columns is `r ncol(hw_problem2_df)`, the mean is `r mean(pull(hw_problem2_df, x))`, the median is `r median(pull(hw_problem2_df, x))`,  the standard deviation is `r sd(pull(hw_problem2_df, x))`, and the proportion of cases for which (x + y > 1) is `r sum(pull(hw_problem2_df, vec_logical), na.rm=TRUE)/500`

_Scatter Plot 1: Logical Variable_
```{r yx_scatter1}
ggplot(hw_problem2_df, aes(x = x, y = y, color = vec_logical)) + geom_point()
ggsave("hw_problem2_df.pdf", width = 4, height = 6)
```
_Scatter Plot 2: Numeric Variable_
```{r yx_scatter2}
ggplot(hw_problem2_df, aes(x = x, y = y, color = vec_numeric)) + geom_point()
```
_Scatter Plot 3: Factor Variable_
```{r yx_scatter3}
ggplot(hw_problem2_df, aes(x = x, y = y, color = vec_factor)) + geom_point()
```

The color scales on these scatterplots indicate the type of variables included. For example, the vec_numeric color scale is a gradient representing continuous variables, the vec_factor and vec_logical variables are distinguished representing categorical data. 


